Reliable feature point matching without the use of geometric constraints

### 2013-07-03 text/clustering.tex
Commited a lot of minor changes

\documentclass[conference]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
%\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

containing viewport changes and a simple method to generate a large 

@@

not fall within paired areas. Examples include \cite{das2008event} and 
\cite{wu2011robust}.

@@

In practice however, there are many situations where using the geometry 
of the image to filter correspondences is not possible, such as object 
recognition where adjacent objects in one photo might be separated in 
another. Another example is images that do not overlap where algorithms 
might still find matches adhering to the imposed geometric constraints.
In addition methods using geometric constraints all require an initial 
set of correspondences. If this set of correspondences is narrowed down 
to a set with a higher ratio of correct matches the result of the 
geometric matching will benefit in terms of speed and accuracy. Finally 
some use cases might require a performance that can't be achieved by a 
more complex geometric method where simple non geometric methods might 
be able to provide sufficient matching accuracy with a smaller 
performance penalty.

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing every 
feature point of one image with all feature points of the other and 
finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. Ranking the 
scores by their uniqueness and picking the $n$ best produces a set of 
correspondences that are distinctly matched across the two images. This 
method will be referred to as \emph{SIFT} in this paper.

@@

%Based on this idea the two proposed methods find reliable matches as 
%follows:
%\begin{itemize}
%\item[]{\emph{Mirror Match (MM)}: Match features using the gold standard 
%		algorithm\cite[p. 114]{multipleView} ranked by similarity and 
%		thresheld by uniqueness\cite{lowe2004sift}.  However instead of 
%		matching features from one image with features in another, we 
%	match every feature with all other features of the two images 
%combined. Only matches from one image to the other are returned.}
%\item[]{\emph{Mirror Match with Clustering (MMC)}: Take the combined set 
%		of feature points from both images and cluster these points 
%		according to their descriptors. Given a resulting partition of 
%		points, no matches are returned if it contains only feature 
%	points from one image. If the partition contains points from both 
%images, \emph{Mirror Match} is used to find the best matches within the 
%partition.}
%\end{itemize}
%%

@@

	\caption{Illustration of matches filtered by \emph{MM}. For each 
	correct match found between the two images, a green line is shown.  
Incorrect matches are marked with a red line. Figure \ref{fig:within} 
shows the correspondences found by mirror match, but including all 
matches that are found within the same image. }%

@@

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping the 2.5\% of 
edges with the highest similarity seems to work well in practice. Figure 
\ref{fig:graph} shows the result of clustering the feature points 
displayed as a graph.
%
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.32\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
		\end{subfigure}%
	}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}

@@

In \cite{mikolajczyk2005performance} Mikolajczyk and Schmid introduces a 
set of images to compare the performance of feature detectors. The test 
set covers different types of image variance such as different lighting, 
blur, rotation as well as viewpoint change. Inspired by this 
dataset\footnote{In particular by the 'Graf' image set} and motivated by 
the need for more image pairs featuring viewport changes, we have 
compiled a set of $8$ image pairs consisting of subjects taken from two 
different angles. The images are collected from Flickr's database of 
images published under a creative commons 
licence\footnote{\href{http://creativecommons.org/}{www.creativecommons.org}, 
a copyleft license permitting redistribution} and feature murals which 
makes it possible to relate points in the image pairs with a homography.  
The images have been cropped to feature the same motive and resized to 
$900$px by $600$px. This dataset will be referred to as the 
\emph{Murals} dataset in this paper. The images can be seen in figure 
\ref{fig:murals}.

@@

%
\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		\end{subfigure}%
		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		\end{subfigure}%
		\enspace %
		\begin{subfigure}[t]{0.36\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/crop_examples}
		\end{subfigure}%
	}%
	\caption{Example of test set produced by a source pair of images}
	\label{fig:fairey}
\end{figure}
%
%\subsection{Experimental setup}
The \emph{MM} and \emph{MMC} algorithms have been tested against 
\emph{SIFT}, the algorithm Lowe proposed in \cite{lowe2004sift} as well 
as the \emph{Isodata} matching algorithm proposed in \cite{das2008event} 
which uses geometric information to improve the matching. The comparison 
has been done using the \emph{Murals} dataset, the \emph{Graf} set from 
\cite{mikolajczyk2005performance} and two images\footnote{100\_1942.jpg 
and 100\_1941.jpg} from \cite{gallagher2008} illustrated in figure 
\ref{fig:comparemirror}.

Test sets have been generated from the image pairs by cropping out 
$250$px by $250$px patches with a random vertical and horizontal offset.  
Given a source image pair we produce $100$ pairs of patches where the 
two patches in any pair might or might not overlap.  This ensures that 
the patch pairs with no overlap still retain a basic similarity to each 
other, while the patches that do overlap often only overlap on a small 
part of their surface. Figure \ref{fig:fairey} shows an example of 
possible pairs of test images produced from a source image pair.  
Producing $n$ such pairs allows us to test not just how well the 
matching algorithm performs on a variety of overlap but also how many 
false positives we get on similar images that do not overlap at 
all\footnote{The set of source images and homographies as well as the 
	script to generate the cropped test sets based on them have been 
	made available online together with the exact test sets used in this 
paper on 
\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.  
In practice the amount of overlap between pairs in a test set will 
depend on the overlap and viewport change of the source image pair.  To 
give a rough impression of a typical example, table \ref{table:overlap} 
shows the overlap of image pairs created from img1 and img3 of the 
\emph{Graf} image set from \cite{mikolajczyk2005performance}.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_viewport}
	\caption{Results on viewport changes using the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} as image pair sources. Img 1 
		to 6 in the \emph{Graph} set are images of the same subject but 
		with gradually increasing viewport change. Img1 and 2 are 
		almost identical while img1 and 6 are taken from very different 
		angles. S is a testset of $100$ crops from img1 and img2. M is 
	likewise but based on img1 and img3.  L is img1 and img4 and XL is 
img1 and img5.  To decrease clutter \emph{MM} is not shown, but it 
performs alike to \emph{MMC}}
	\label{fig:result_viewport}
\end{figure}

Figure \ref{fig:result_graf} shows the results for 100 image pairs 
generated from img1 and img3 of the \emph{Graf} image set. The plot 
illustrates the results of \emph{SIFT}, \emph{MM}, \emph{MMC} and 
\emph{Isodata} as thresholds are varied over a range. \emph{Isodata} 
performs well with strict thresholds, but when we require more matches 
\emph{MM} and \emph{MMC} show superior results. To see if this results 
generalizes over more image sets, we tested the four algorithms on the 
\emph{Murals} dataset as well as the \emph{Graf} pair tested in fig.  
\ref{fig:result_graf}. In total 900 image pairs were generated from 9 
source pairs generating the plot shown in figure 
\ref{fig:result_accumulated} displaying the accumulated amount of 
matches on the x-axis. To verify that this performance is independent of 
the magnitude of viewport change, we tested \emph{MMC} and \emph{SIFT} 
on the \emph{Graf} image set as shown in figure 
\ref{fig:result_viewport}. The figure confirms that \emph{MMC} is 
generally superior across viewport changes.

For an example of a real life use case, figure \ref{fig:result_pitts} 
shows the results on 100 potentially overlapping image pairs generated 
from a typical holiday scene featuring occlusion and a slight viewport 
change.

@@

\section{Summary}
We have addressed the problem of matching feature points without using 
geometrical constraints and proposed two algorithms \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. The two algorithms 
share the same common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches and \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points. We 
also address how to test the algorithms by constructing the 
\emph{Murals} dataset consisting of $8$ source image pairs and a way to 
generate testsets based on these pairs.  The proposed approach shows 
promising results on both the \emph{Murals} dataset as well as a real 
life example from the Gallagher dataset and the \emph{Graf} image set, 
surpassing even a method using geometrical constraints when matching 
image pairs that might not overlap.
%
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}

>>>>

\documentclass[conference]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
%\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
%\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

containing viewpoint changes and a simple method to generate a large 

@@

not fall within paired areas. Examples include \cite{das2008event} which 
clusters the feature points according to their position in the image
and \cite{wu2011robust} which uses points returned by MSER feature 
points to designate areas. In this paper the algorithm from  
\cite{das2008event} will later be referred to as \emph{Isodata}.

@@

In practice there are many situations where using the geometry of the 
image to filter correspondences is not possible, such as object 
recognition where adjacent objects in one photo might be separated in 
another. Another example is images that do not overlap where algorithms 
might still find matches adhering to the imposed geometric constraints.
In addition methods using geometric constraints all require an initial 
set of correspondences. If this set of correspondences is narrowed down 
to a set with a higher ratio of correct matches the result of the 
geometric matching will benefit in terms of speed and accuracy. Finally 
some use cases might require a performance that can't be achieved by a 
more complex geometric method where simple non geometric methods might 
be able to provide sufficient matching accuracy with a smaller 
performance penalty.

Without using geometric constraints we are left with a few methods to 
match feature points. As pointed out in \cite{szeliski2010} the simplest 
approach to matching feature point is to return all matches within a 
threshold of similarity.  This can be done with a fixed threshold or by 
finding the nearest neighbors for each point, still filtering by a 
threshold to avoid including matches for feature points that have no 
counterpart in the other image.  In \cite{lowe2004sift}, Lowe proposes 
an alternative measure where the uniqueness of a given match is found by 
looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities.  Ranking the 
scores by their uniqueness and picking the $n$ best produces a set of 
correspondences that are distinctly matched across the two images. This 
method will be referred to as \emph{SIFT} in this paper. Of these three 
methods\footnote{fixed threshold, nearest neighbors and \emph{SIFT}}, 
\emph{SIFT} was found to perform best by a slight margin in a study in 
\cite{mikolajczyk2005performance}.

@@

Based on this idea the two proposed methods find reliable matches as 
follows. In \emph{Mirror Match (MM)} we match every feature from both 
images with all other features and filter matches based on the ratio of 
best to second best match as in \emph{SIFT}. In \emph{Mirror Match with 
Clustering (MMC)} on the other hand we take the combined set of feature 
points from both images and cluster these points according to their 
descriptors.  Given a resulting partition of points, no matches are 
returned if it contains only feature points from one image.  If the 
partition contains points from both images, \emph{Mirror Match} is used 
to find the best matches within the partition.

@@

	\caption{Illustration of matches filtered by \emph{MM}. a) is the 
		source image pair. b) displays the feature points found in the 
		two images. c) shows the result of \emph{SIFT} matching the two 
		images. For each correct match found between the two images, a 
		green line is shown.  Incorrect matches are marked with a red 
		line. d) depicts all matches found by \emph{MM} including 
	matches between points in the same image. e) shows the result from 
d) but with the intra-image matches removed. f) shows an example of a 
partition of feature points from the \emph{MMC} algorithm.}%

@@

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping the 2.5\% of 
edges with the highest similarity seems to work well in practice. Figure 
\ref{fig:graph} shows the result of clustering the feature points 
displayed as a graph.
%
\begin{figure}
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
		\includegraphics[width=\textwidth]{images/MMC_graph_full}
		\caption{Full Graph}
		\label{fig:full_graph}
	\end{subfigure}%

	%add desired spacing between images, e. g. ~, \quad, \qquad		  
	%(or a blank line to force the subfigure onto a new line)
	\begin{subfigure}[t]{0.36\textwidth}
		\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
		\caption{Cropped Section of Graph}
		\label{fig:cropped_graph}
	\end{subfigure}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{figure}

@@

In \cite{mikolajczyk2005performance} Mikolajczyk and Schmid introduces a 
set of images to compare the performance of feature detectors. The test 
set covers different types of image variance such as different lighting, 
blur, rotation as well as viewpoint change. Inspired by this 
dataset\footnote{In particular by the 'Graf' image set} and motivated by 
the need for more image pairs featuring viewpoint changes, we have 
compiled a set of $8$ image pairs consisting of subjects taken from two 
different angles. The images are collected from Flickr's database of 
images published under a creative commons 
licence\footnote{\href{http://creativecommons.org/}{www.creativecommons.org}, 
a copyleft license permitting redistribution} and feature murals which 
makes it possible to relate points in the image pairs with a homography.  
The images have been cropped to feature the same motive and resized to 
$900$px by $600$px. This dataset will be referred to as the 
\emph{Murals} dataset in this paper. The images can be seen in figure 
\ref{fig:murals}.

\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		\end{subfigure}%
		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		\end{subfigure}%
		\enspace %
		\begin{subfigure}[t]{0.36\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/crop_examples}
		\end{subfigure}%
	}%
	\caption{Example of test set produced by a source pair of images}
	\label{fig:fairey}
\end{figure}

@@

%
%
%\subsection{Experimental setup}
The \emph{MM} and \emph{MMC} algorithms have been compared to 
\emph{SIFT}, the algorithm Lowe proposed in \cite{lowe2004sift} as well 
as the \emph{Isodata} matching algorithm proposed in \cite{das2008event} 
which uses geometric information to improve the matching. The comparison 
with \emph{SIFT} serves to illustrate the relative performance of 
\emph{MM} and \emph{MMC} to a state of the art algorithm not using 
geometric constraints. \emph{Isodata} is included in the comparison to 
illustrate the performance of an algorithm relying on geometric 
constraints faced with image pairs that have might not have any overlap.  
The comparison has been done using the \emph{Murals} dataset, the 
\emph{Graf} set from \cite{mikolajczyk2005performance} and two 
images\footnote{100\_1942.jpg and 100\_1941.jpg} from 
\cite{gallagher2008} illustrated in figure \ref{fig:comparemirror}.

Test sets have been generated from the image pairs by cropping out 
$250$px by $250$px patches with a random vertical and horizontal offset.  
Given a source image pair we produce $100$ pairs of patches where the 
two patches in any pair might or might not overlap.  This ensures that 
the patch pairs with no overlap still retain a basic similarity to each 
other, while the patches that do overlap often only overlap on a small 
part of their surface. Figure \ref{fig:fairey} shows an example of 
possible pairs of test images produced from a source image pair.  
Producing $n$ such pairs allows us to test not just how well the 
matching algorithm performs on a variety of overlap but also how many 
false positives we get on similar images that do not overlap at 
all\footnote{The set of source images and homographies as well as the 
script to generate the cropped test sets based on them will be made 
available online together with the exact test sets used in this paper 
upon acceptance}
%\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.  
In practice the amount of overlap between pairs in a test set will 
depend on the overlap and viewpoint change of the source image pair.  To 
give a rough impression of a typical example, table \ref{table:overlap} 
shows the overlap of image pairs created from img1 and img3 of the 
\emph{Graf} image set from \cite{mikolajczyk2005performance}.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_viewpoint}
	\caption{Results on viewpoint changes using the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} as image pair sources. Img 1 
		to 6 in the \emph{Graph} set are images of the same subject but 
		with gradually increasing viewpoint change. Img1 and 2 are 
		almost identical while img1 and 6 are taken from very different 
		angles. S is a testset of $100$ crops from img1 and img2. M is 
	likewise but based on img1 and img3.  L is img1 and img4 and XL is 
img1 and img5.  To decrease clutter \emph{MM} is not shown, but it 
performs alike to \emph{MMC}}
	\label{fig:result_viewpoint}
\end{figure}

Figure \ref{fig:result_graf} shows the results for 100 image pairs 
generated from img1 and img3 of the \emph{Graf} image set. The plot 
illustrates the results of \emph{SIFT}, \emph{MM}, \emph{MMC} and 
\emph{Isodata} as thresholds are varied over a range. \emph{Isodata} 
performs well with strict thresholds, but when we require more matches 
\emph{MM} and \emph{MMC} show superior results. To see if this results 
generalizes over more image sets, we tested the four algorithms on the 
\emph{Murals} dataset as well as the \emph{Graf} pair tested in fig.  
\ref{fig:result_graf}. In total 900 image pairs were generated from 9 
source pairs generating the plot shown in figure 
\ref{fig:result_accumulated} displaying the accumulated amount of 
matches on the x-axis. To verify that this performance is independent of 
the magnitude of viewpoint change, we tested \emph{MMC} and \emph{SIFT} 
on the \emph{Graf} image set as shown in figure 
\ref{fig:result_viewpoint}. The figure confirms that \emph{MMC} is 
generally superior across viewpoint changes.

For an example of a real life use case, figure \ref{fig:result_pitts} 
shows the results on 100 potentially overlapping image pairs generated 
from a typical holiday scene featuring occlusion and a slight viewpoint 
change.

@@

\section{Summary}
We have addressed the problem of matching feature points without using 
geometrical constraints and proposed two algorithms \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}.  The two algorithms 
share the same common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches and \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points. The 
algorithms show promising results when tested on the \emph{Murals} data 
set where pairs of the same object seen from different angles are 
cropped to have different degrees of overlap ranging from full to non at 
all. We show that this result generalizes to variations in viewpoint 
change by comparing \emph{MMC} to \emph{SIFT} over a set of image pairs 
with an increasing magnitude of perspective difference. We go on to 
apply the algorithms to a real life image case featuring occlusion and a 
slight viewpoint change and show that the performance of \emph{MM} and 
\emph{MMC} is consistent with the results on \emph{Murals}. Given the 
versatility of the proposed algorithms, it would be interesting to apply 
them to problems that require high reliability faced with images that 
might not match, such as near duplicate detection or face clustering.
%
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}
